Update of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-11-29 DryDep.tex
Fixed some formulations to make things clearer.

\subsubsection*{Mapping of land surface types}
Land surface types from, e.g., \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM}) 2 categories, have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}). For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From the MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert. This mapping differs from the one used in the Wesely scheme.

@@

  \caption{Mapping of land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM) 2 to the nine land surface types used in the EMAP scheme. Water bodies and snow and ice categories of MODIS are actually not mapped. For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From MODIS category 'Barren or sparsely vegetated', everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with modeling results}
\label{subsec:model}
%
In our evaluation of results, we closely follow suggestions by \citet{ACP:Hardacre2015}. For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-gridded our data to a horizontal resolution of $3^\circ\times 3^\circ$. In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our sensitivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types which have been retrospectively separated (Section~\ref{subsubsec:seasons}). From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare also with results from \citet{ACP:Luhar2017}.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistency with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low absolute variance. Most remarkable but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference (at most $3.5\,\unit{\%}$ increase in the deep tropics). Switching from the SMOKE-BEIS parameterization to the pre-processed greening season (\emph{EMEP\_ppgs} compared to \emph{EMEP\_full}), surface ozone increases on average by $6\,\unit{\%}$ between $54-70\,\unit{^\circ N}$. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, probably owing to the industrial development of countries in these regions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conclusions[Discussion and conclusions]  %% \conclusions[modified heading if necessary]
\label{sec:conc}
We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere into account \citep{ACP:Simpson2012}.The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans. Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average. However, there seems to be a tendency, that the newer TF~HTAP-models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models are decreasing ozone dry deposition \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.

Most of this decrease is due to changes in the corresponding dry deposition velocities at the order of one magnitude in case of desert and grasslands, or due to the accumulative effect of small changes over the ocean which is especially important in the southern hemisphere. With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition over the former and overestimate it over the latter. Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}, or influence of desert dust and wind drag.

@@

We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics is most sensitive to changes in ozone uptake by plants. This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics. Tropospheric total column ozone in the tropics is predicted to decrease \citep[e.g.,][]{WMO2014}, while tropospheric and surface ozone increases. The combined effects of increasing emissions of ozone precursors and an increase in UV due to thinning of stratospheric ozone might permit more UV light at ground and thus increase the ozone production.

@@

Future work on the OsloCTM3 should include a direct output of dry deposition velocities for diagnostic purposes. The model performance could also be improved by allowing for more plant fictionally types than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE). The photolysis- and chemical reaction computation as well as reaction rates shall also be revised in the future.

>>>>

\subsubsection*{Mapping of land surface types}
Land surface types from, e.g., \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM})~2 categories, have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}). For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From the MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert. This mapping differs from the one used in the Wesely scheme.

@@

  \caption{Mapping of land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM)~2 to the nine land surface types used in the EMAP scheme. Water bodies and snow and ice categories of MODIS are actually not mapped. For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From MODIS category 'Barren or sparsely vegetated', everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with modeling results}
\label{subsec:model}
%
In our evaluation of results, we closely follow suggestions by \citet{ACP:Hardacre2015}. For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF~HTAP) models, we also have re-gridded our data to a horizontal resolution of $3^\circ\times 3^\circ$. In Section~\ref{subsubsec:zonal}, we look at zonal distributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our sensitivity simulations and study seasonal cycles of hemispheric ozone as well as for nine land surface types which have been retrospectively separated (Section~\ref{subsubsec:seasons}). From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare also with results from \citet{ACP:Luhar2017}.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistency with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low absolute variance. Most remarkable but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference (at most $3.5\,\unit{\%}$ increase in the deep tropics). Switching from the SMOKE-BEIS parameterization to the pre-processed greening season (\emph{EMEP\_ppgs} compared to \emph{EMEP\_full}), surface ozone increases on average by $6\,\unit{\%}$ between $54-70\,\unit{^\circ N}$. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, probably owing to the industrial development of countries in these regions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conclusions[Discussion and conclusions]  %% \conclusions[modified heading if necessary]
\label{sec:conc}
We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere into account \citep{ACP:Simpson2012}. The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans. Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average. However, there seems to be a tendency, that the newer TF~HTAP-models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models are decreasing ozone dry deposition \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.

Most of this decrease in the OsloCTM3 is due to changes in the corresponding dry deposition velocities at the order of one magnitude in case of desert and grasslands, or due to the accumulative effect of a small change in dry deposition to the ocean which is especially amplified in the southern hemisphere. With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the new OsloCTM3 dry deposition scheme slightly underestimates ozone dry deposition velocities over the former and overestimate it over the latter. Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile to study dry deposition in these regimes at a more process-oriented level, e.g., 2-layer gas exchange with ocean waters \citep{ACP:Luhar2017}, wave braking and spray \citep{ACP:Pozzer2006}, or influence of desert dust.

@@

We have studied the parameter space of the stomatal conductance parameterization and found that total surface ozone in the tropics is most sensitive to changes in ozone uptake by plants. This may indicate that future changes in vegetation cover and solar radiation at the surface due to changes in stratospheric ozone, cloud cover, or aerosols could also strongly influence the surface ozone burden in the tropics. Total column ozone in the tropics is predicted to decrease \citep[e.g.,][]{WMO2014}, while tropospheric and surface ozone increase. The combined effects of increasing emissions of ozone precursors and an increase in UV due to thinning of stratospheric ozone might permit more UV light at ground and thus increase the ozone production.

@@

Future work on the OsloCTM3 should include a direct output of dry deposition velocities for diagnostic purposes. The model performance could also be improved by allowing for more plant fuctional types than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE). The photolysis- and chemical reaction computation as well as reaction rates shall also be revised in the future.
